---
title: "First Gen Students: A Different Experience"
format: revealjs
editor: visual
---

## The Data

-   Provided by Macalester's Institutional Research office
-   Data regarding previous graduated students from Fall 2013 to Spring 2018
-   Approximately 2500 total students

## The Sample {.smaller}

```{r, include=FALSE, echo=FALSE}
library(readr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
majors <- read_csv("https://raw.githubusercontent.com/wastjohn/COMP112-Final/main/majordata.csv")
fgen <- majors %>% filter(is.na(first_gen) == FALSE)
```

```{r, echo=FALSE}
ggplot(majors, aes(x=first_gen, fill = sex)) + #add fill = sex maybe?
  geom_bar() +
  facet_grid(~grad_year) +
  scale_fill_viridis_d() +
  labs(title = 'First Gen Distribution by Sex from Fall 2013 to Spring 2018', x = 'Type of Student', y = 'Count', fill = 'Sex')
```

-   Approximately 120 first gen graduates
-   Not a lot of first gen data till 2016
-   Very specific sample
-   Drop the NA (need to do this)

## The Sample: Limited

```{r}
ggplot(fgen, aes(x=first_gen, fill = sex)) +
  geom_bar() + #add position = fill for a prop bar chart
  facet_grid(~grad_year) +
  scale_fill_viridis_d() +
  labs(title = 'First Gen Distribution by Sex from Fall 2015 to Spring 2018', x = 'Type of Student', y = 'Count', fill = 'Sex')
```

-   Overall More Females than Males

## First Gen Majors: Major 1

```{r}
ggplot(fgen %>% filter(major_ct == 1), aes(x = major_ct, fill = major1_division)) +
  geom_bar(position = 'fill') +
  facet_grid(~first_gen)+
  scale_fill_viridis_d() +
  labs(title = 'Proportion of First Major Division', x = 'Number of Majors', y = 'Percent', fill = 'Division of First Major') +
  scale_x_discrete(limits = 1)
```

## First Gen Majors: Major 2

```{r}
ggplot(fgen %>% filter(major_ct == 2), aes(x = major_ct, fill = major2_division)) +
  geom_bar(position = 'fill') +
  facet_grid(~first_gen)+
  scale_fill_viridis_d() +
  labs(title = 'Proportion of Second Major Division', x = 'Number of Majors', y = 'Percent', fill = 'Division of Second Major') +
  scale_x_discrete(limits = 2)
```

## First Gen Majors: Major 3

```{r}
ggplot(fgen %>% filter(major_ct == 3), aes(x = major_ct, fill = major3_division)) +
  geom_bar(position = 'fill') +
  facet_grid(~first_gen)+
  scale_fill_viridis_d() +
  labs(title = 'Proportion of Third Major Division', x = 'Number of Majors', y = 'Percent', fill = 'Division of Third Major') +
  scale_x_discrete(limits = 3)
```

## First Major: Fine Arts

```{r}
ggplot(fgen %>% filter(major1_division == 'Fine Arts'), aes(x = first_gen, fill = major1)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Fine Arts Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Fine Arts First Major', x = 'Type of Student', y = 'Percent', fill = 'Fine Arts Major')
```

## First Major: Humanities

```{r}
ggplot(fgen %>% filter(major1_division == 'Humanities'), aes(x = first_gen, fill = major1)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Humanities Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Humanities First Major', x = 'Type of Student', y = 'Percent', fill = 'Humanities Major')

```

## First Major: Interdisciplinary

```{r}
ggplot(fgen %>% filter(major1_division == 'Interdisciplinary'), aes(x = first_gen, fill = major1)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Interdisciplinary Major Distribution')+ 
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Interdisciplinary First Major', x = 'Type of Student', y = 'Percent', fill = 'Interdisciplinary Major')

```

## First Major: Natural Sciences and Mathematics

```{r}
ggplot(fgen %>% filter(major1_division == 'Natural Sciences and Mathematics'), aes(x = first_gen, fill = major1)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Natural Sciences and Mathematics Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Natural Sciences and Mathematics First Major', x = 'Type of Student', y = 'Percent', fill = 'Natural Sciences and Mathematics Major')
```

## First Major: Social Sciences

```{r}
ggplot(fgen %>% filter(major1_division == 'Social Sciences'), aes(x = first_gen, fill = major1)) +
  geom_bar(position = 'fill') +
  scale_fill_viridis_d() +
  labs(title = 'Social Sciences Major Distribution')+
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Social Sciences First Major', x = 'Type of Student', y = 'Percent', fill = 'Social Sciences Major')
```

## Second Major: Fine Arts

```{r}
ggplot(fgen %>% filter(major2_division == 'Fine Arts'), aes(x = first_gen, fill = major2)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Fine Arts Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Fine Arts Second Major', x = 'Type of Student', y = 'Percent', fill = 'Fine Arts Major')
```

## Second Major: Humanities

```{r}
ggplot(fgen %>% filter(major2_division == 'Humanities'), aes(x = first_gen, fill = major2)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Humanities Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Humanities Second Major', x = 'Type of Student', y = 'Percent', fill = 'Humanities Major')
```

## Second Major: Interdisciplinary

```{r}
ggplot(fgen %>% filter(major2_division == 'Interdisciplinary'), aes(x = first_gen, fill = major2)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d()+
  labs(title = 'Interdisciplinary Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Interdisciplinary Second Major', x = 'Type of Student', y = 'Percent', fill = 'Interdisciplinary Major')
```

## Second Major: Natural Sciences and Mathematics

```{r}
ggplot(fgen %>% filter(major2_division == 'Natural Sciences and Mathematics'), aes(x = first_gen, fill = major2)) +
  geom_bar(position = 'Fill') +
  scale_fill_viridis_d() +
  labs(title = 'Natural Science and Mathematics Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Natural Sciences and Mathematics Second Major', x = 'Type of Student', y = 'Percent', fill = 'Natural Sciences and Mathematics Major')
```

## Second Major: Social Sciences

```{r}

ggplot(fgen %>% filter(major2_division == 'Social Sciences'), aes(x = first_gen, fill = major2)) +
  geom_bar(position = 'fill') +
  scale_fill_viridis_d() +
  labs(title = 'Social Sciences Major Distribution') +
  facet_wrap(~sex) +
  labs(title = 'Proportion of Students with Social Sciences Second Major', x = 'Type of Student', y = 'Percent', fill = 'Social Sciences Major')
```

## Studied Away By Sex

```{r}
ggplot(fgen, aes(x = studied_away, fill = sex)) +
  geom_bar(position = 'Fill') +
  facet_grid(~first_gen) +
  scale_fill_viridis_d() +
  labs(title = 'Proportion of Students Who Studied Abroad', x = 'Stayed Home or Studied Abroad', y = 'Percent', fill = 'Sex') +
  scale_x_discrete(limits = c(0, 1), labels = c('Home', 'Abroad'))
```

## Conclusions {.scrollable, .smaller}
-   No Triple Majors
-   No Males in Fine Arts
-   No Theater or Dance Majors
-   No Females in Physics
-   No Males Anthropology or Psychology
-   Males Like Econ
-   Low Selection Diversity Overall in Second Major (except Natural Sciences and Mathematics and Females in Social Sciences)
-   No Females in Russian or Religous Studies
-   No Women/Gender Studies
-   Biology popular in Females


## Limitations/Assumptions {.scrollable}
-   Small sample size
-   Fewer first gen students overall
-   First Gen records begin later
-   More females than males
-   Forms were filled out correctly

